\mytitle

### 2023-05-13 chapters/concl.tex
Complete concl & impl todos

There is still more work to be done in support of this thesis.
% TODO better transition sentence
%In addition to the cases presented, there are many more classes of systems
%that could benefit from program synthesis tools.
One interesting direction for work in this area is expanding on \jitsynth
to allow for some class of compiler optimizations.
In order to further improve the runtime performance of synthesized compilers,
\jitsynth could use superoptimization techniques to minimize the size
of discovered minicompilers.
Exploring how to best apply these techniques to improve \jitsynth's effectiveness
would be an interesting direction for future work.
%One interesting direction for work in this area is
%automating the implementation of blockchain staking protocols with program synthesis.
%Existing work has developed techniques for synthesizing more traditional distributed protocols, \todo{cite transit}
%and properties for staking protocols are specified in a fairly straightforward mannor,
%but complex required invariants on ___ make implementing staking mechanisms difficult.
%Automating this implementation step could prevent bugs in these protocols
%and the massive loss of funds that comes with these exploits.
Another promising direction lies in further improving \depsynth's usability.
Existing techniques for litmus test generation and partial interpretation of litmus tests
\todo{cite crashmonkey and memsynth}
could allow \depsynth to synthesize dependency rules more quickly.
% Distributed protocol synthesis: https://www.cis.upenn.edu/~alur/udupa-thesis.pdf
Finally, my overall thesis would be further supported by a larger
body of synthesis tools for a wide array of modern systems.
\cref{c:depsynth} discusses existing works in synthesis for
techniques for both memory models and distributed protocols,
but other promising system domains for program synthesis %application
include blockchain staking protocols and security monitors for IFC kernels.
%automating the implementation of blockchain staking protocols with program synthesis.
%Existing work has developed techniques for synthesizing more traditional distributed protocols, \todo{cite transit}
Security and liveness properties for staking protocols are specified in a fairly straightforward manner,
but complex invariants on the blockchain state make implementing staking mechanisms difficult.
\todo{cite https://academic.oup.com/rfs/article-abstract/34/3/1156/5868423, https://link.springer.com/chapter/10.1007/978-3-319-67816-0_17, https://link.springer.com/chapter/10.1007/978-3-319-63688-7_12}
Automating this implementation step could prevent bugs in staking protocols
and the massive loss of funds that comes with these exploits.
Information flow control (IFC) kernels give developers the ability to enforce
a variety of security properties by limiting the information flow between applications.
However, correctly implementing these restrictions is prohibitively difficult.
Existing work enables push-button verification of these security properties for IFC kernels,\todo{cite}
but still requires a manually-written implementation.
By synthesizing security monitors for IFC kernels that guarantee these properties,
the high implementation burden for IFC security could be removed entirely.
Overall, program synthesis tools can help systems developers build correct, performant software
more effectively and easily for a wide spectrum of system domains.

>>>>

There is still more work to be done in support of this thesis.
% TODO better transition sentence
%In addition to the cases presented, there are many more classes of systems
%that could benefit from program synthesis tools.
One interesting direction for work in this area is expanding on \jitsynth
to allow for some class of compiler optimizations.
In order to further improve the runtime performance of synthesized compilers,
\jitsynth could use superoptimization techniques to minimize the size
of discovered minicompilers.
Exploring how to best apply these techniques to improve \jitsynth's effectiveness
would be an interesting direction for future work.
%One interesting direction for work in this area is
%automating the implementation of blockchain staking protocols with program synthesis.
%Existing work has developed techniques for synthesizing more traditional distributed protocols, \todo{cite transit}
%and properties for staking protocols are specified in a fairly straightforward mannor,
%but complex required invariants on ___ make implementing staking mechanisms difficult.
%Automating this implementation step could prevent bugs in these protocols
%and the massive loss of funds that comes with these exploits.
Another promising direction lies in further improving \depsynth's usability.
Existing techniques for litmus test generation and partial interpretation of litmus tests
\cite{mohan:crashmonkey,bornholt:memsynth}
could allow \depsynth to synthesize dependency rules more quickly.
% Distributed protocol synthesis: https://www.cis.upenn.edu/~alur/udupa-thesis.pdf
Finally, my overall thesis would be further supported by a larger
body of synthesis tools for a wide array of modern systems.
\cref{c:depsynth} discusses existing works in synthesis for
techniques for both memory models and distributed protocols,
but other promising system domains for program synthesis %application
include blockchain staking protocols and security monitors for IFC kernels.
%automating the implementation of blockchain staking protocols with program synthesis.
%Existing work has developed techniques for synthesizing more traditional distributed protocols, \todo{cite transit}
Security and liveness properties for staking protocols are specified in a fairly straightforward manner,
but complex invariants on the blockchain state make implementing staking mechanisms difficult
\cite{fahad:pos,wenting:secure-pos,aggelos:verified-pos}.
Automating this implementation step could prevent bugs in staking protocols
and the massive loss of funds that comes with these exploits.
Information flow control (IFC) kernels give developers the ability to enforce
a variety of security properties by limiting the information flow between applications.
However, correctly implementing these restrictions is prohibitively difficult.
Existing work enables push-button verification of these security properties for IFC kernels \cite{sigurbjarnarson:nickel},
but still requires a manually-written implementation.
By synthesizing security monitors for IFC kernels that guarantee these properties,
the high implementation burden for IFC security could be removed entirely.
Overall, program synthesis tools can help systems developers build correct, performant software
more effectively and easily for a wide spectrum of system domains.

### 2023-05-13 chapters/depsynth/impl.tex
Complete concl & impl todos

\paragraph{Attempting graph search with a \tord solver.}
One bottleneck in the time \depsynth takes to find a complete set of
dependency rules comes from the time it takes to search for a minimal
happens-before graph for a given litmus test.
This section explores a strategy that we explored to limit this bottleneck.
Specifically, we encode the happens-before graph search problem
in the theory of ordering consistency (\tord)\todo{cite}
and execute the search with an existing \tord solver.
Ultimately, we found that this different search strategy did not 
improve the overall runtime, but we find this negative
result worth reporting.

% TODO elaborate on \tord

% TODO present results
Intuitively, % TODO discuss why this didn't work

>>>>

\subsection{Attempting graph search with a \tord solver.}
One bottleneck \depsynth
comes from the time it takes to search for a minimal
happens-before graph for a given litmus test.
This section describes a strategy that we explored to limit this bottleneck.
Specifically, we encode the happens-before graph search problem
in the theory of ordering consistency ($\tord$) \cite{n:tord}
and execute the search with an existing $\tord$ solver.
Ultimately, we found that this different search strategy did not 
improve the overall runtime, but we find this negative
result worth reporting.

% TODO present results
As described in \cref{sec:alg:onetest}, the \rulesfortest algorithm
finds dependency rules for a litmus test by searching for a total order over writes,
reducing the total order to a minimal happens-before graph,
then generalizing that graph into rules.
The reduction to graph and generalization into rules steps are both straightforward,
but since the space of possible total orders grows exponentially with the number of writes,
finding valid orders over writes can be exceedingly slow for large tests.
While the pruning techniques presenting in \cref{sec:alg}
and the heuristics over prioritizing program order help mitigate the runtime cost of this search,
there remains space for further improvement.

In an attempt to speed up the overall search time for total orders over writes,
we encode the problem in $\tord$ and offloading to a $\tord$ solver.
$\tord$ defines predicates for describing partial-ordering constraints
on reads and writes for multi-threaded programs, % TODO
but can also be used to describe orderings of storage system writes.
Specifically, $\tord$ introduces the predicate $\prec$, where
$ w_i \prec w_j $ encodes the fact that write $w_i$ occured before
write $w_j$.
To encode our search in $\tord$,
we describe the existence of each happens-before graph edge
between write events $e_i$ and $e_j$ as the boolean variable $d_{ij}$,
then introduce the following constraint: $ d_{ij} \rightarrow e_i \prec e_j $.
By solving these constraints along with the user-defined constraints specifying
crash consistency for the storage system, we arrive at a total order over all
writes for the given litmus test.

Existing work defines a DPLL(T)-style solver for $\tord$ \cite{n:tord},
which we use for solving the constraints specified above and
determining a total order over writes for single litmus tests.
However, by doing so, we empirically found that the solver
slightly diminished the performance of \depsynth for all three of our evaluated case studies
(in the case of \shardstore, slowing the rules search from
\shardstoresynthesistime minutes to \shardstoresynthesistimetord minutes).
Intuitively, we believe this performance degredation is an artifact of 
bad initial assignments from the SAT solver
and a resulting a large number of iterations in the DPLL(T) loop.
For storage systems with increasingly complex behavior,
\depsynth may require much larger litmus tests,
which may make the benefits of searching for orders with $\tord$ more apparent.
However, for the case studies considered in this paper,
we found our original search algorithm from \cref{sec:alg:onetest}
to outperform the $\tord$-directed search.
